---
title: "distribution/regression"
author: "Jieqi Tu"
date: "2/1/2021"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Import datafile
```{r import data}
biopsy.grade = readxl::read_excel("./data/GG_AP_ClinicalOutcomes_Biostats.xlsx", sheet = "Biopsy Grade Group")
adverse.pathology = readxl::read_excel("./data/GG_AP_ClinicalOutcomes_Biostats.xlsx", sheet = "Adverse Pathology")
serum.geneglobe = readxl::read_excel("./data/Serum_SerumEV_GeneGlobeRQ_CT35.xlsx", sheet = "Serum_GeneGlobeRQ_CT35")
EV.geneglobe = readxl::read_excel("./data/Serum_SerumEV_GeneGlobeRQ_CT35.xlsx", sheet = "EV_GeneGlobeRQ_CT35")
```

## Tidy Data
```{r tidy data}
# change format from wide to long
# serum
new = t(serum.geneglobe)
name_tmp <- new[1,]
new <- new[-1,]
new <- apply(new,1,as.numeric) %>% t()
colnames(new) <- (name_tmp)
serum.geneglobe.new = new %>% as.data.frame()
serum.geneglobe.new = rownames_to_column(serum.geneglobe.new)
colnames(serum.geneglobe.new)[1] = "Patient ID"

# EV
new = t(EV.geneglobe)
name_tmp <- new[1,]
new <- new[-1,]
new <- apply(new,1,as.numeric) %>% t()
colnames(new) <- (name_tmp)
EV.geneglobe.new = new %>% as.data.frame()
EV.geneglobe.new = rownames_to_column(EV.geneglobe.new)
colnames(EV.geneglobe.new)[1] = "Patient ID"

# Adverse pathology
adverse.pathology$`Adverse Pathology` = ifelse(adverse.pathology$`Adverse Pathology` == "99", NA, adverse.pathology$`Adverse Pathology`)
```

## Checking distribution for each microRNA
```{r distribution}
# checking distribution for serum
normality.test.serum = data.frame(
  microRNA = colnames(serum.geneglobe.new[-1]),
  p.value = numeric(61)
)

for (i in 1:61) {
  test = shapiro.test(serum.geneglobe.new[,i+1])
  normality.test.serum$p.value[i] = test$p.value
}

# checking distribution for EV
normality.test.EV = data.frame(
  microRNA = colnames(EV.geneglobe.new[-1]),
  p.value = numeric(61)
)

for (i in 1:61) {
  test = shapiro.test(EV.geneglobe.new[,i+1])
  normality.test.EV$p.value[i] = test$p.value
}

# Results display
normality.test.serum
normality.test.EV
```

Conclusion: None of the microRNA has the normal distribution. 

## Univariate GLM fitting

#### Biopsy Grade
```{r datafile combine}
# datafile combine
# left join for grade groups
serum.grade = left_join(serum.geneglobe.new, biopsy.grade, by = "Patient ID")
EV.grade = left_join(EV.geneglobe.new, biopsy.grade, by = "Patient ID")

# left join for adverse pathology
serum.all = left_join(serum.grade, adverse.pathology, by = "Patient ID")
EV.all = left_join(EV.grade, adverse.pathology, by = "Patient ID")
```

```{r Univariate GLM}
# univariate GLM
grade.serum.result = data.frame(
  microRNA = colnames(serum.geneglobe.new)[-1],
  p.value = numeric(61)
)
for (i in 1:61) {
  model = glm(formula = serum.all$`Biopsy Grade Group`~serum.all[,i+1])
  result = summary(model)
  grade.serum.result$p.value[i] = result$coefficients[2,4]
}

grade.EV.result = data.frame(
  microRNA = colnames(EV.geneglobe.new)[-1],
  p.value = numeric(61)
)

for (i in 1:61) {
  model = glm(formula = EV.all$`Biopsy Grade Group`~EV.all[,i+1])
  result = summary(model)
  grade.EV.result$p.value[i] = result$coefficients[2,4]
}

# Display result
grade.serum.result
grade.EV.result
```

Now we want to focus on the significant predictors.
```{r model selection}
grade.serum.result.significant = 
  grade.serum.result %>% filter(p.value <= 0.05)

grade.EV.result.significant = 
  grade.EV.result %>% filter(p.value <= 0.05)

# backward selection for serum/grade model
library(MASS)
fullmodel1 = glm(data = serum.all, formula = `Biopsy Grade Group`~`hsa-miR-1246`+`hsa-miR-214-3p`+`hsa-miR-301a-3p`)
serum.grade.final = stepAIC(object = fullmodel1, direction = "backward")
```

Therefore the final model only contains hsa-miR-1246 and hsa-miR-214-3p for serum/biopsy grade.
The final model only contains hsa-miR-31-5p for EV/biopsy grade.

#### Adverse Pathology
```{r GLM for adverse pathology}
# univariate GLM
adverse.serum.result = data.frame(
  microRNA = colnames(serum.geneglobe.new)[-1],
  p.value = numeric(61)
)
for (i in 1:61) {
  model = glm(formula = serum.all$`Adverse Pathology`~serum.all[,i+1], family = "binomial")
  result = summary(model)
  adverse.serum.result$p.value[i] = result$coefficients[2,4]
}

adverse.EV.result = data.frame(
  microRNA = colnames(EV.geneglobe.new)[-1],
  p.value = numeric(61)
)

for (i in 1:61) {
  model = glm(formula = EV.all$`Adverse Pathology`~EV.all[,i+1], family = "binomial")
  result = summary(model)
  adverse.EV.result$p.value[i] = result$coefficients[2,4]
}

# Display result
adverse.serum.result
adverse.EV.result
```

Now we want to focus on the significant predictors.
```{r model selection2}
adverse.serum.result.significant = 
  adverse.serum.result %>% filter(p.value <= 0.05)

adverse.EV.result.significant = 
  adverse.EV.result %>% filter(p.value <= 0.05)

# backward selection for EV/adverse pathology model
library(MASS)
fullmodel2 = glm(data = EV.all, formula = `Adverse Pathology`~`hsa-let-7a-5p`+`hsa-miR-26b-5p`+`hsa-miR-301a-3p`, family = "binomial")
EV.adverse.final = stepAIC(object = fullmodel2, direction = "backward")
```

The final model only contains hsa-let-7a-5p for EV/adverse pathology.
