---
title: "Univariate GLM for Biopsy Grade"
author: "Jieqi Tu"
date: "3/15/2021"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Method

Biopsy grade groups were dichotomized as "Low" (if <=2) and "High" (if > 2).
Univariate generalized linear models were used to detect association between dichotomous biopsy grades and all the serum and serum microRNAs. The logit link function was used when modeling. All the significance level was set as 0.05.

```{r import dataset, include=FALSE}
biopsy.grade = readxl::read_excel(path = "./data/3-1 data/GG_AP_ClinicalOutcomes_Biostats_Updated.xlsx", sheet = "Biopsy Grade Group")
biopsy.grade$grades = ifelse(biopsy.grade$`Biopsy Grade Group` <= 2, "low", "high") %>% as.factor()
biopsy.grade = biopsy.grade[,-2]
serum = readxl::read_excel(path = "./data/Serum_AdjustedRQ_CT35_BioStats.xlsx", sheet = "geNorm NormFinder RQ")
EV = readxl::read_excel(path = "./data/SerumEV_AdjustedRQ_CT35_BioStats.xlsx", sheet = "geNorm NormFinder RQ")
```



```{r data manipulation, include=FALSE}
# Change the format

# Serum
new = t(serum)
name_tmp <- new[1,]
new <- new[-1,]
new <- apply(new,1,as.numeric) %>% t()
colnames(new) <- (name_tmp)
serum = new %>% as.data.frame()
serum = rownames_to_column(serum)
colnames(serum)[1] = "Patient ID"

# EV
new = t(EV)
name_tmp <- new[1,]
new <- new[-1,]
new <- apply(new,1,as.numeric) %>% t()
colnames(new) <- (name_tmp)
EV = new %>% as.data.frame()
EV = rownames_to_column(EV)
colnames(EV)[1] = "Patient ID"
```



```{r combine dataset, include=FALSE}
serum.new = left_join(serum, biopsy.grade, by = "Patient ID")
EV.new = left_join(EV, biopsy.grade, by = "Patient ID")
```

## Results for Univariate Generalized Linear Models

Since hsa-miR-130b-3p only have value 0, here we excluded this variable.

```{r Univariate GLM, warning=FALSE}
EV.new = EV.new[,-7]

serum.result = data.frame(
  microRNA = colnames(serum.new)[-c(1,63)],
  coefficient.estimates = numeric(61),
  standard.deviation = numeric(61),
  p.value = numeric(61)
)


for (i in 1:61) {
  model = glm(formula = serum.new$`grades`~serum.new[,i+1], family = "binomial")
  result = summary(model)
  serum.result$coefficient.estimates[i] = result$coefficients[2,1]
  serum.result$standard.deviation[i] = result$coefficients[2,2]
  serum.result$p.value[i] = result$coefficients[2,4]
}

EV.result = data.frame(
  microRNA = colnames(EV.new)[-c(1,62)],
  coefficient.estimates = numeric(60),
  standard.deviation = numeric(60),
  p.value = numeric(60)
)

for (i in 1:60) {
  model = glm(formula = EV.new$`grades`~EV.new[,i+1], family = "binomial")
  result = summary(model)
  EV.result$coefficient.estimates[i] = result$coefficients[2,1]
  EV.result$standard.deviation[i] = result$coefficients[2,2]
  EV.result$p.value[i] = result$coefficients[2,4]
}

# Show the results
serum.result = serum.result[order(serum.result$p.value),]; serum.result
EV.result = EV.result[order(EV.result$p.value),]; EV.result

# Print significant variables
serum.result %>% 
  filter(p.value <= 0.05)

EV.result %>% 
  filter(p.value <= 0.05)
```

## Variable selection 

We now first look into the model fitting 
```{r variable selection for serum miRNAs}
# Variable selection for serum miRNAs
library(MASS)
fullmodel.serum = glm(grades~`hsa-miR-18b-5p`+`hsa-miR-1246` +`hsa-miR-363-3p` + `hsa-miR-16-5p` + `hsa-miR-320a` + `hsa-miR-301a-3p` + `hsa-miR-27a-3p` + `hsa-miR-146a-5p`, family = "binomial", data = serum.new)
serum.step = stepAIC(object = fullmodel.serum, direction = "backward")
finalmodel.serum = glm(grades~`hsa-miR-18b-5p` +`hsa-miR-363-3p` + `hsa-miR-320a` + `hsa-miR-301a-3p`, family = "binomial", data = serum.new)
summary(finalmodel.serum)

# Variable selection for EM serum miRNAs
fullmodel.EV = glm(grades~`hsa-miR-19b-3p` + `hsa-miR-451a` + `hsa-miR-143-3p` + `hsa-miR-25-3p` + `hsa-miR-210-3p` + `hsa-miR-874-3p` + 
                     `hsa-miR-363-3p` + `hsa-miR-18b-5p` + `hsa-miR-222-3p` + `hsa-miR-30a-5p`, family = "binomial", data = serum.new)
EV.step = stepAIC(object = fullmodel.EV, direction = "backward")
finalmodel.EV = glm(grades~`hsa-miR-18b-5p`  + `hsa-miR-25-3p` + `hsa-miR-363-3p`, family = "binomial", data = EV.new)
summary(finalmodel.EV)
```


## Classification
```{r classification}
library(caret)
library(pROC)
set.seed(1029)
rowTrain = createDataPartition(y = serum.new$grades,
                               p = 0.75,
                               list = FALSE)

ctrl = trainControl(method = "repeatedcv",
                    repeats = 5,
                    summaryFunction = twoClassSummary,
                    classProbs = TRUE)

# Get column numbers for best predictors for serum miRNAs
which(colnames(serum.new)=="hsa-miR-363-3p")
which(colnames(serum.new)=="hsa-miR-18b-5p")
which(colnames(serum.new)=="hsa-miR-301a-3p")
which(colnames(serum.new)=="hsa-miR-320a")

# Logistic regression
glm.serum = train(x = serum.new[rowTrain, c(37, 20, 31, 32)],
                                  y = serum.new$grades[rowTrain],
                                  method = "glm",
                                  metric = "ROC",
                                  trControl = ctrl)

# See the prediction performance
glm.predict = predict(glm.serum, newdata = serum.new[-rowTrain,], type = "prob")[,2]
roc.glm.serum = roc(serum.new$grades[-rowTrain], glm.predict)
plot(roc.glm.serum, legacy.axes = T, print.auc = T)

# Get column numbers for best predictors for EV serum miRNAs
which(colnames(EV.new)=="hsa-miR-18b-5p")
which(colnames(EV.new)=="hsa-miR-25-3p")
which(colnames(EV.new)=="hsa-miR-363-3p")

# Logistic regression
rowTrain = createDataPartition(y = EV.new$grades,
                               p = 0.75,
                               list = FALSE)
glm.EV = train(x = EV.new[rowTrain, c(19, 55, 36)],
                                  y = EV.new$grades[rowTrain],
                                  method = "glm",
                                  metric = "ROC",
                                  trControl = ctrl)

# See the prediction performance
glm.predict.EV = predict(glm.EV, newdata = EV.new[-rowTrain,], type = "prob")[,2]
roc.glm.EV = roc(EV.new$grades[-rowTrain], glm.predict.EV)
plot(roc.glm.EV, legacy.axes = T, print.auc = T)
```

